---
title: "UE20CS312 - Data Analytics - Worksheet 1b - Correlation Analysis"
subtitle: "PES University"
author:
- 'Purvik S Nukal, Dept. of CSE - PES1UG20CS315'
output: pdf_document
urlcolor: blue
editor_options:
markdown:
wrap: 72
---

```{r}
library(ggpubr)

```


```{r}
library(dplyr)

```

```{r}
library(ltm)

```

```{r}
library(psych)
```

```{r}
df <- read.csv('D:/PES University/5th sem/Data analytics/lab1b/road_accidents_india_2016.csv', row.names=1)
```


#**Problem1**


```{r}
acc_cols <- grep("Total.accidents", colnames(df), ignore.case=T, value = TRUE)
```


```{r}
totalAccidents <- data.frame(state..ut=df$State..UT, total_acc=rowSums(df[ ,(acc_cols)], na.rm = TRUE))
```


```{r}
print.data.frame(totalAccidents)
```


```{r}
death_cols <- grep("Persons.Killed$", colnames(df), ignore.case = T, value = TRUE)
```


```{r}
totalAccidents$total_deaths <- rowSums(df[ ,c(death_cols)])
```


```{r}
totalAccidents$fatality_rate <- totalAccidents$total_deaths/totalAccidents$total_acc
```


```{r}
totalAccidents$mist_rate <- df$Mist..Foggy...Total.Accidents/totalAccidents$total_acc
```


```{r}
print.data.frame(totalAccidents)
```



#**Problem2**


```{r}
plot(x = totalAccidents$fatality_rate, y = totalAccidents$mist_rate)
```


```{r}
 r <- cor(totalAccidents$fatality_rate, totalAccidents$mist_rate)
```


```{r}
r
```


```{r}
corr_test <- cor.test(totalAccidents$fatality_rate, totalAccidents$mist_rate, method = 'pearson')
```


```{r}
print(corr_test)

```

From the given outputs we can conclude that since the p value (being 0.08231) is greater than the significance level of 0.05 so we cannot reject the NULL hypothesis.Hence we do not have sufficient evidence to conclude that there is a significant linear relationship between the *fatality rate* of a state and the *mist/foggy* rate.

#**Problem3**


```{r}
totalAccidents$acc_ranks <- rank(desc(totalAccidents$total_acc), ties.method = 'random')
```


```{r}
totalAccidents$death_ranks <- rank(desc(totalAccidents$total_deaths), ties.method = 'random')
```


```{r}
rs <- cor(totalAccidents$acc_ranks, totalAccidents$death_ranks, method = 'spearman')
```

```{r}
rs
```


```{r}
degrees <- nrow(totalAccidents)-2
```


```{r}
t_stat <- (rs - 0.2)/sqrt((1-rs*rs)/(nrow(totalAccidents)-2))
```


```{r}
t_stat
```

```{r}
2*pt(q=t_stat, df=degrees, lower.tail = FALSE)
```

From the following we can conclude that when tested for a correlation of at least 0.2 with a significance level of 0.05, the t statistic value came out to be 15.1203 as a result of which we rejected the NULL hypothesis and and concluded that there is a strong linear correlation between the 2 rank columns.


#**Problem4**

```{r}
df$hao <- ifelse(df$Hail.Sleet...Total.Accidents>0,1,0)
a = biserial.cor(y = df$hao, x = df$Rainy...Total.Accidents, use = c("complete.obs"))
a
```

#**Problem5**

```{r}
df$dao <- ifelse(df$Dust.Storm...Total.Accidents>0,1,0)
ndf = data.frame("Dust_Accidents"= df$dao,"Hail_Accidents"=df$hao)
contingency = table(ndf)
contingency
```

```{r}
phi(contingency, digits = 10)
```